---
title: "Policy Cost"
format:
  html:
    fig-width: 8
    fig-height: 6
---


# Scoping the Cost

Utilizing the current active employee population, we need to calculate the following items:

* The current base compensation (I have convert the hourly workers to year compensation)
* New compensation with 6% raise
* Burden rate (estimated between 25% and 40% of employee's base salary)

Because the compensation is at a yearly value and if we assume the cost of living increase would go in place of the merit, this would the give us our projected cost for 2026.

```{r, char-init}
#| echo: false
#| warning: false
#| error: false
library(tidyverse)

source(here::here("src", "_initial_settings.R"))

emp_clean <- readRDS(file = here::here("data", "employee.RDS")) |> 
  filter(!is.na(Term_Date)) 

```



# Analysis Overview

```{r}
#| echo: true
#| warning: false
#| error: false
#| code-fold: true

emp_clean |>  # filtered down to only active employees
  summarise(TotalCompensation = sum(EffectiveSalary, na.rm=T)) |> 
  ungroup() |> 
  mutate(NewComp = TotalCompensation * 1.06) |> 
  mutate(IncreaseDuetoRaise = NewComp - TotalCompensation) |> 
  mutate(Burden_25 = IncreaseDuetoRaise * 0.25,
          Burden_40 = IncreaseDuetoRaise * 0.4) |> 
  mutate(TotalCost_25 = Burden_25 + IncreaseDuetoRaise,
          TotalCost_40 = Burden_40 + IncreaseDuetoRaise) |> 
  pivot_longer(cols = 1:7) |>   
  flextable()

```


Therefore to raise everyone's pay by 6% the total cost would be between **$11.6M - $13M** 

* The cost for 6% raises is $9.3M
* The additional employeer cost due to payroll taxes / benfits: between $2.3M and $3.7M

