---
title: "Policy Cost"
format:
  html:
    fig-width: 8
    fig-height: 6
---


::: {.callout-important collapse="true"}
## Additional Analysis: Correlation to Voluntary

Formula for total cost To accurately budget for the increase, calculate the following for the entire company or for the employees receiving the raise:\((\text{Total\ Current\ Base\ Salary}\times 1.06)\times (1+\text{Burden\ Rate})-(\text{Total\ Current\ Base\ Salary}\times (1+\text{Burden\ Rate}))\)The "Burden Rate" represents the cost of employer-paid benefits and taxes, usually estimated between 25% and 40% of an employee's base salary. 

:::

# Analysis Overview

Utilizing the current active employee population, we need to calculate the following items:

* The current base compensation (I have convert the hourly workers to year compensation)
* New compensation with 6% raise
* Burden rate (estimated between 25% and 40% of employee's base salary)

```{r, char-init}
#| echo: false
#| warning: false
#| error: false
library(tidyverse)

source(here::here("src", "_initial_settings.R"))

emp_clean <- readRDS(file = here::here("data", "employee.RDS")) |> 
  filter(!is.na(Term_Date)) 

```



# Analysis Overview

```{r}
#| echo: true
#| warning: false
#| error: false
#| code-fold: true

emp_clean |> 
  summarise(TotalCompensation = sum(EffectiveSalary, na.rm=T)) |> 
  ungroup() |> 
  mutate(NewComp = TotalCompensation * 1.06) |> 
  mutate(IncreaseDuetoRaise = NewComp - TotalCompensation) |> 
  mutate(Burden_25 = IncreaseDuetoRaise * 0.25,
          Burden_40 = IncreaseDuetoRaise * 0.4) |> 
  mutate(TotalCost_25 = Burden_25 + IncreaseDuetoRaise,
          TotalCost_40 = Burden_40 + IncreaseDuetoRaise) |> 
  pivot_longer(cols = 1:7) |>   
  flextable()

```


Therefore to raise everyone's pay by 6% the total cost would be between **$11.6M - $13M** 

* The cost for 6% raises is $9.3M
* The additional employeer cost due to payroll taxes / benfits: between $2.3M and $3.7M

