---
title: "Turnover"
format:
  html:
    fig-width: 10
    fig-height: 6
---

```{r, init}
#| echo: false
#| warning: false
#| error: false
library(tidyverse)

source(here::here("src", "_initial_settings.R"))
source(here::here("src", "monthly_employee.R"))

emp_clean <- readRDS(file = here::here("data", "employee.RDS"))
emp_dates <- readRDS(file = here::here("data", "employee_dates.RDS"))

```

::: {.callout-important collapse="true"}
## Additional Analysis: Increase in Voluntary in Sept 2025

There is a slight increase in the voluntary turnover in Sept 2025, would want to dig further into this to determine if this would seem to be a trend or if there is something else going on here.

From the BLS data, it looks like the market supports that there is a delta between voluntary and involuntary, but it doesn't seem like there's a trend up for voluntary recently.

:::


# Trend of Exits 

To start our understanding of employee turnover trends we are first looking at the monthly exits trended from 2018 to now and then we have calculated the year turnover rates. 

* We see the trend of exits (as well as yearly turnover rates) saw a significant change from pre-pandemic levels during both 2022 and 2023.
* While 2025 YTD turnover higher than pre-pandemic levels, the average monthly levels have significantly decreased.

```{r, exit-trend}
#| echo: false
#| warning: false
#| error: false
#| fig-height: 4
#| fig-align: center
source(here::here("src", "trend_exits_turnover.R"))

  pw2
```


# Macro-economic Trends

We want to understand how Jackson compares to the larger economic forces in order to assess if Jackson turnover generally aligns. 

* We do see that the job openings have slowed in the Finance and Insurance industry since the start of 2024, and the quit rate was higher in 2022 and 2023 than before or after that time frame.
   * This would indicate that the Finance industry nationwide allows had a similar trend of a "Great Resignation" in 2022-2023 and has generally moved to the "Great Stay" as new openings have cooled.

```{r}
#| echo: false
#| warning: false
#| error: false
#| out-width: "100%"
#| fig-height: 4
#| fig-align: center

source(here::here("src", "jolts_data.R"))

 bls_p1 +
    scale_x_date(date_breaks = "1 year", date_labels = "%Y") + 
  scale_color_manual(values = jcolor) +
  labs(x = NULL, y= NULL, color = NULL) + 
  theme(
    panel.background = element_rect(fill = "#F8F5F5"),
    plot.background = element_rect(fill = "#F8F5F5"),
    legend.background = element_rect(fill = "#F8F5F5"),
    legend.key = element_rect(fill = "#F8F5F5", color = "darkgrey")
  )

```

# Digging into Turnovers

## Voluntary vs Involuntary

Interestingly we see that both voluntary and involuntary turnover increases saw spikes in 2022 and 2023, with the levels going back down in 2024 and 2025. 

* This would seem to indicate that there were both significant layoffs in these two years as well as significant amount of employees who voluntarily quit. 

```{r}
#| echo: false
#| warning: false
#| error: false
#| out-width: "100%"
#| fig-height: 4
#| fig-align: center

turnover_month <- emp_dates |> 
  # head(10) |> 
  rename(start_date = Hire_Date,
        end_date = CleanEnd) |> 
  select(employee_id = ID, start_date, end_date) |> 
  create_monthly_activity() |> 
  left_join( , y= emp_clean |> select(ID), 
      by = join_by(employee_id == ID)) |> 
  group_by(month) |> 
  summarise(active_employee = n()) |> 
  ungroup() |> 
  filter(month >= as.Date("2018-01-01")) |> 
  left_join( , y= emp_dates |> 
                  left_join(., y=emp_clean |> filter(!is.na(Term_Type)) |> select(ID, Term_Type), by = join_by(ID)) |> 
                  group_by(Term_Month, Term_Type) |> 
                  summarise(exits = n()) |> 
                  ungroup() |> 
                  filter(!is.na(Term_Type)), 
        by = join_by(month == Term_Month)) |> 
  mutate(exits = ifelse(is.na(exits), 0, exits)) |> 
  mutate(turnover = exits/active_employee) |> 
  ungroup() |> 
  filter(!is.na(Term_Type))


turnover_month |> 
  filter(month < as.Date("2025-10-01")) |> 
    filter(month >= as.Date("2018-01-01")) |> 
    ggplot(aes(month, turnover, color = Term_Type))+
    geom_rect(aes(xmin = as.Date("2022-01-01"), xmax = as.Date("2023-01-01"), ymin = -Inf, ymax = Inf), 
      alpha = 0.01, fill = jcolor[6])+
    geom_rect(aes(xmin = as.Date("2023-01-01"), xmax = as.Date("2024-01-01"), ymin = -Inf, ymax = Inf), 
      alpha = 0.01, fill = jcolor[5])+
    geom_point()+
    geom_line()+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") + 
  scale_color_manual(values = jcolor) +
  scale_y_continuous(labels = scales::percent_format())+
  labs(title = "Count of Employees Exits By Month and Termination Type",
    x = NULL, y= NULL) + 
  theme(
    panel.background = element_rect(fill = "#F8F5F5"),
    plot.background = element_rect(fill = "#F8F5F5"),
    legend.background = element_rect(fill = "#F8F5F5"),
    legend.key = element_rect(fill = "#F8F5F5", color = "darkgrey")
  )
```



## Notes
I built "monthly_employee.R" as a function that could be called to get the active number of employees per month given a dataset like this HRIS sample. This would allow for a quick recreated of monthly view when we are testing future policy impacts.